DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2017-01-04 2015CompleteDerivation/completeDerivation.tex
Cleanup and added biblio items

While some branches of physics (e.g. Newtonian mechanics, thermodynamics, special relativity) are derived from physical principles (e.g. Newton's laws, laws of thermodynamics, invariance of the speed of light), others start by simply setting a mathematical framework. It is the case for classical Lagrangian mechanics, where we start by assuming the existence of a Lagrangian function whose path integral is minimized during the motion. It is the case for classical Hamiltonian mechanics, where we start by assuming the state is described by conjugate pairs $(q^i, p_i)$ and evolves according to Hamilton's equations. It is the case for quantum mechanics, where the state space is a complex inner product vector space and the state evolves according to Schroedinger's equations. That is, at the very core of some of our physical theories we have mathematical definitions instead of physical ideas.

If we ask ``when can a particular physical system be suitably described by Hamiltonian (or by classical) mechanics?" the only answer we have is ``when it fits the mathematical framework". While many seem to be content with this answer (e.g. ``you have to start from somewhere, it may as well be the math which at least is precise"), it begs the question: it does not tell us how and what physical properties of the system under study are being captured.

This is a problem since mathematical structures are not enough to characterize the physical system being described. For example, the same math for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems. The mathematical treatment is the same because it captures aspects of the description that are common to all cases. But for the very same reason, we cannot infer what system is being described if all we are given is the mathematical model. In the jargon of linguistic/computer science/phylosophy: math captures the syntax (the relationships between the objects) but does not capture the semantic (the meaning of each object).

We started this work with the unpretentious goal of exploring simple questions such as the following. What do the mathematical symbols used in classical/quantum Hamiltonian represent? What is the physical meaning attached to it? Why is the state of a quantum particle described by a wave function (i.e. a vector in a complex inner product space), while the state of a classical particle is described by the pairs $(q^i, p_i)$ (i.e. a point in a cotangent bundle)? The idea was to create a math-to-physics dictionary so that each mathematical concept was clearly mapped to a crisp physical concept. But to make sure that such dictionary was complete, we realized we had to make sure that the whole relevant mathematical framework could be recovered.

This redefined our mission to re-deriving as much as possible starting from physical assumptions. It also forced us to push further and further back toward the basic definitions. For example, if we truly wanted to justify why the state space of Hamiltonian mechanics is charted by conjugate variables $q^i$ and $p_i$, we had to justify why it was a manifold as well, which in turn meant we needed to justify why it was a topological space. But when we did find physical concepts that mapped to these more basic mathematical definitions, we realized that this link was valid for a far wider range of cases: the concepts were more general than we anticipated. For example, we realized that the mathematical notion of a topology maps incredibly well to the idea of physical distinguishability, but states of Hamiltonian system are not the only thing that are physically distinguishable.

Each new link often shrank the complexity of the arguments while expanding the scope of the work in a way that was not at all anticipated. The same link that allowed to justify Hamiltonian dynamics also gave us relativistic motion on the extended phase space. The same link that allowed to justify Lagrangian mechanics also constrained the motion to that of free particles under scalar/vector potential forces. We also realized that sometimes different mathematical concepts could be justified from the same physical starting point. That is: by putting physics first we could achieve a sort of unification of ideas that was not practicable by math alone.

@@

While trying to re-derive classical, quantum and (to some extent) relativistic mechanics may seem too broad a scope, it is necessary to show that the concepts we develop are truly general, that they can be applied to many different cases. If we could derive classical Hamiltonian mechanics on assumptions that are inapplicable to the quantum case, what good would it do? Quantum system are Hamiltonian systems, so clearly we wouldn't have understood what makes a system Hamiltonian. The ability to apply the same physical ideas far and wide is what gives us confidence that these are truly general concept that capture the physical aspects we are describing. The breath itself is what makes the work interesting and more than a curiosity.

@@

%TODO: Point out most modern physics starts by assuming some mathematical framework. Pros and cons of that.

%TODO: High level view of the derivation

%While some branches of fundamental physics (e.g. Newtonian mechanics and special relativity) are founded on physical laws or principles others (e.g. Lagrangian/Hamiltonian/Quantum mechanics) start by postulating mathematical frameworks or relationships. The latter approach presents a significant drawback: the mathematical structure is not enough to define the physical system it describes. We are left asking: under what physical conditions a particular system is described by Lagrangian (or Quantum) mechanics? Naturally one \emph{can} define a Lagrangian system (or a Quantum system) as one that obeys those rules, but that begs the question. Another issue is that each element of the mathematical structure may or may not correspond to some physical concept. Suppose the earning (or losses) of two companies are governed by Hamiltonian's equations, which variable is on the cotangent space of the other? In other words: if we want to fully appreciate what the fundamental theories of physics are describing, what each mathematical concept corresponds to in the physical world, we need to start by characterizing the physical system we are describing in a precise enough way that it can be encoded in mathematical language. Naturally, the physical insight we gain will be relevant for only that type of system (i.e. other different systems may \emph{happen} to use the same equations), but what we'll see is that the assumptions we make are minimal and can apply to a variety of systems, especially the ones that are considered fundamental. It is indeed surprising how so much can be derived by imposing so little.

%This work aims to re-organize the known elements and equations in a more consistent and comprehensive way, leading to better insight on why the fundamental concepts and laws are what they are. We will start with physical assumptions, that clarify the models we impose on the physical world, and give arguments on when those assumptions can be considered valid. We will justify our mathematical definitions, based on the formal properties of the objects of our discussion. And will, as a consequence, re-derive known results and theories. We will strive to do this in a way that is mathematically meaningful, philosophically consistent and mathematically precise.

%We'll use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics, and we'll find interesting connections among them. We'll keep names and notation as consistent as possible to current use across the different disciplines. This may sometimes lead to some non sequitur as it will not be immediately clear why the new definitions are equivalent to the standard ones. These are typically resolved by subsequent derivation of the expected properties.

%No mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible. No proof is longer than a couple of paragraphs, so the word \emph{theorem} is avoided in favor of \emph{proposition} and \emph{corollary}. The novel, and surprising, result is how so much can be derived from so little.

@@

The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology, measure theory, differential geometry (symplectic and Rimannian), group theory and vector spaces, statistics, information theory, Hamiltonian mechanics, Lagrangian mechanics, quantum mechanics, thermodynamics, special relativity and so on. Using idea from all these areas is necessary, in fact unraveling their connections is part of the value of this work, but it does create a number of problems.

@@

This link between state definition and deterministic processes should not be too surprising as the state, in the context of thermodynamics and systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}. As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can indeed be calculated, measured and prepared. We can also describe the evolution of each element of the distribution provided that: we have a way to isolate it and study it under deterministic and reversible motion (so that we can define microstates); the non-deterministic motion does not alter the system (the set of microstates is preserved by the evolution).

@@

However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy is finite.} Note that different outcomes for the same process can overlap. For example, $4.12 \pm 0.05$ cm and $4.13 \pm 0.05$ cm are both legitimate possible outputs of the same measurement device. But since all states must be distinguishable, given two arbitrary states there must be a process precise enough to tell them apart. That is, the potential outcomes associated with the two states do not overlap. For example, $4.12 \pm 0.0005$ cm and $4.13 \pm 0.0005$ cm do not overlap anymore.

@@

Given that scientific reductionism (i.e. the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics, we now explore how to characterize a system in terms of its components. That is, we want to study the relationship between the state space of a composite system and the state space of its parts. In general, this is quite a complicated thing to do, which requires intimate knowledge of the system at hand. So we simplify our problem and study a material made of infinitesimal homogeneous parts. What we'll find is that under the additional assumption that the material is infinitesimally reducible (i.e. its state is equivalent to the states of its infinitesimal parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.

@@

As we identified the space of Lebesgue integrable differentiable functions $C^1(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the space of distributions that are physically meaningful, we can better understand why other commonly used function spaces do not fit the bill. Some are not restrictive enough. The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{prop:continuity} are unphysical. The space of continuous functions $C(\mathcal{S})$, the space of continuous functions that vanish at infinity $C_0(\mathcal{S})$ and the space of differentiable functions $C^1(\mathcal{S})$ include functions whose integral is infinite, which would represent an infinite amount of material. Some definitions are too restrictive. Requiring compact support (i.e. the function is different from zero only on a finite region) would exclude distributions, such as Gaussians, that span over the whole range of states. Such distributions are physical: if we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range. Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible. The Whitney approximation theorem makes this mathematically well defined.}

@@

But probably the easiest way to look at it is that initial and final sets of states have to possess the same count of states and possibilities as defined by the symplectic form $\omega$. Therefore the area within a degree of freedom, which represents the number of possibilities in said d.o.f., needs to be mapped to an equal area within the transported d.o.f. (i.e. the d.o.f. defined by the transported state variables). Independent d.o.f. must remain independent, and therefore transported orthogonal d.o.f. remain orthogonal. This means that the product of possibilities of independent d.o.f. is also conserved. These statements are the physical justification of Gromov's non-squeezing theorem and Liouville's theorem. And they give intuitive insight on the geometry of Hamiltonian systems.

@@

The third case is when we perform a state variable transformation $\hat{q}=\hat{q}(q,t)$ that is time dependent (e.g. $\hat{q}=q+vt$). Our composite state distribution $\rho(q)$ becomes $\rho(\hat{q},t)$: it is no longer defined at equal time. This means that our measure $\mu$ needs to be modified to define integrals over the time variable.\footnote{Looking ahead, another way to frame the problem is that Galilean transformations do not preserve phase space volume. Lorentz transformations, on the other hand, do and it is this requirements plus the kinematic assumption that will recover them later.}

@@

As we said before, the time variable is not simply another state variable: it doesn't identify extra configurations and it's the variable on which we have defined deterministic and reversible evolution. When changing variables, we still need to make sure that deterministic and reversible motion can be defined on the new time variable. Transformations like $\hat{t}=q, \; \hat{q}=t$ (exchanging time with another state variable) or $\hat{t}=\sqrt{t^2 + q^2}, \; \hat{q}=\tan^{-1}(t/q)$ (polar coordinates between time and state variable) clearly do not guarantee deterministic and reversible motion over the new time variable. To be meaningful, the change of time variable must preserve time ordering therefore the transformation has to be strictly monotonic between the two time variables. On the other hand, the change of state variables must preserve the ability to uniquely identify states at each instant in time. In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight.}

@@

@@

The link between kinematics, the study of motion through quantities like velocity and acceleration, and dynamics, the study of state evolution and its causes, is crucial in physics. In Newtonian physics, $\vec{F}=m\vec{a}$ provides such link. In Hamiltonian (and Lagrangian) mechanics, the link is implicitly given by the choice of the Hamiltonian (or Lagrangian).

@@

This also allows us to understand conceptually why the speed of deterministic and reversible motion cannot be greater than $c$. Consider the trajectory of a point particle. Over an interval $dt$ it will travel $dx$ in space. Since the motion is deterministic and reversible, and therefore continuous, it will travel through all the states defined in the interval $dx$. It will also travel through every moment in time, which corresponds to $cdt$ possibilities. As it moves, time will always change but not necessarily space. The number of states traveled in space cannot exceed the number of states traveled in time. Therefore we have:
\begin{align*}
\frac{|n_x(dx)|}{|n_t(dt)|} &=  \frac{|dx|}{|c dt|} \leq 1 \\
\frac{|dx|}{|dt|} & \leq c \\
\end{align*}
Note, though, that this constraint applies only to deterministic and reversible motion. It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$. Non-local quantum effects, such as the ones featured in EPR experiments, are not examples of deterministic and reversible motion. Therefore they do not have anything to violate: the maximum speed does not apply to them. On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.

@@

Suppose we have our classical electron described by a finite amount of classical material distributed over a single d.o.f. $\int \rho_\mathcal{c} dq \wedge dp = 1$ where $\rho_\mathcal{c}$ is the normalized distribution. In this context, we will call \emph{fragment} an infinitesimal part of our classical electron. The Shannon entropy associated with the distribution is finite as well. $I(\rho_\mathcal{c}) = - \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$. Intuitively, this value can be thought of as the amount of information required to identify a fragment within the distribution.\footnote{To be precise, Shannon's differential entropy is actually the number of bits required to identify an element of the given distribution relative to one of a uniform distribution of unitary range on all axes.}

@@

	This notion may seem in contradiction to the current practice. For example, the S-matrix, one of the main devices for calculating scattering amplitudes and cross sections, is derived by assuming unitary evolution. Doesn't it imply we are assuming that deterministic and reversible evolution is happening at every instant?

@@

%Note on composite systems and distinguishability. When putting together two classical system, just sum distributions. When putting together quantum system we have to clarify: combine to create one quantum system (sum in vector space) or two separate quantum system (symmetrized tensor product). In classical mechanics, 2a + 2b = (a+b) + (a+b). In quantum note |a> x |b> = |a>|b> + |b>|a> but (|a> + |b>) x (|a> + |b>) = |a>|a> + ... That is two quantum states in two different states is not the same as two quantum states spread equally over those two states. In classical they are.)

% Muon example. Consider a muon and its decay into an electron and two neutrinos: it is clear that the three outgoing particles have a state and trajectory of their own, it is clear that the resulting total mass and energy came from the muon. Yet, before the decay, we cannot ascribe an internal independent state and evolution to each part. The state of a muon is not some combination of the state of an electron and two neutrinos. That is: the unstated part is not just microstates.

%TODO: moved from classical
%This intuitive picture is unfortunately not suitable to be generalized to different contexts, so we develop another.

%Consider the following expression:
%\begin{align*}
%\mathcal{c} = \sum\limits_{i \in \mathbbm{I}} a_i %\mathcal{e}(i)
%\end{align*}
%$\mathbbm{I}$ represents all the possible values of a state variable representing the best description of an infinitesimal part that a composite system can give. $\mathcal{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$. $a_i$ represents a transformation that changes the whole state, without affecting the value $i$ for any infinitesimal part. For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables (i.e. the state vector); the only transformation left is increasing/decreasing the amount of particles by a scalar multiple. For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ also represents an internal transformation of such ensemble. This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes how the parts are composed. Giving a full description of the state space of a decomposable system means fully characterizing these components: the state variable $\mathbbm{I}$ and the set of transformations $A$. In the classical case $\mathcal{e}(i)$ form a basis of a real vector space, where $i$ is the state vector for an infinitesimal part, and $a_i$ a real valued coefficient that represents the amount of particles in each state, which describes how the parts are combined.

@@

\bibitem{Shannon} Shannon, C. E., ``A mathematical theory of communication'', The Bell System Technical Journal, Vol. 27, pp. 379--423, 623--656, (1948).
\bibitem{Jaynes} Jaynes, E. T., ``Information theory and statistical mechanics'', Statistical Physics 3, pp. 181--218, (1963).
\bibitem{classical_dynamics} J. V. Jos\'{e}, E. J. Saletan, ``Classical Dynamics'', Cambridge University Press, (1998).
\bibitem{Gromov} Gromov, M. L., ``Pseudo holomorphic curves in symplectic manifolds'', Inventiones Mathematicae 82, pp. 307--347, (1985).
\bibitem{deGosson} de Gosson, M. A., ``The symplectic camel and the uncertainty principle: the tip of an iceberg?'', Foundations of Physics 39, pp. 194--214, (2009).
\bibitem{Stewart} Stewart, I., ``The symplectic camel'', Nature 329, pp. 17--18, (1987).
\bibitem{Lanczos} Lanczos, C., ``The variational principles of mechanics'', University of Toronto Press, (1949).
\bibitem{Synge} Synge, J. L., ``Classical dynamics'', Encyclopedia of Physics Vol 3/1, Springer (1960).
\bibitem{Struckmeier} Struckmeier, J., ``Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems'', J. Phys. A: Math. Gen. 38, pp. 1257--1278, (2005).

>>>>

While some branches of physics (e.g. Newtonian mechanics(ref?), thermodynamics(ref?), special relativity(ref?)) are derived from physical principles (e.g. Newton's laws, laws of thermodynamics, invariance of the speed of light), others start by simply setting a mathematical framework. It is the case for classical Lagrangian mechanics(ref?), where we start by assuming the existence of a Lagrangian function whose path integral is minimized during the motion. It is the case for classical Hamiltonian mechanics(ref?), where we start by assuming the state is described by conjugate pairs $(q^i, p_i)$ and evolves according to Hamilton's equations. It is the case for quantum mechanics(ref?), where the state space is a complex inner product vector space and the state evolves according to Schroedinger's equations. That is, at the very core of some of our physical theories we have mathematical definitions instead of physical ideas.

If we ask ``when can a particular physical system be suitably described by classical Hamiltonian mechanics?" the only answer we have is ``when it fits the mathematical framework". While many seem to be content with this answer (e.g. ``you have to start from somewhere, it may as well be the math which at least is precise"), it begs the question: it does not tell us how and what physical properties of the system under study are being captured.

This is a problem since mathematical structures are not enough to characterize the physical system being described. For example, the same mathematical framework for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems.\cite{Roland,Barron,Borutzky} The mathematical treatment is the same because it captures aspects of the description that are common to all cases. But for the very same reason, we cannot infer what system is being described if all we are given is the mathematical model. In the jargon of linguistic/computer science/phylosophy\cite{Freiding,Friedman}: math captures the syntax (the relationships between the objects) but does not capture the semantic (the meaning of each object).

We started this work with the unpretentious goal of exploring simple questions such as the following. What do the mathematical symbols used in classical/quantum Hamiltonian represent? What is the physical meaning attached to it? Why is the state of a quantum particle described by a wave function (i.e. a vector in a complex inner product space), while the state of a classical particle is described by the pairs $(q^i, p_i)$ (i.e. a point in a cotangent bundle)? The idea was to create a math-to-physics dictionary so that each mathematical concept was clearly mapped to a crisp physical concept. But to make sure that such dictionary was complete, we realized we had to make sure that the whole relevant mathematical framework could be recovered: the translation needs to work both ways.

This redefined our mission to re-deriving as much as possible of the mathematical structure by starting from the smallest set of physical assumptions. It also forced us to push further and further back toward the basic definitions. For example, if we truly wanted to justify why the state space of Hamiltonian mechanics is charted by conjugate variables $q^i$ and $p_i$, we had to justify why it was a manifold as well, which in turn meant we needed to justify why it was a topological space. But when we did find physical concepts that mapped to these more basic mathematical definitions, we realized that this link was valid for a far wider range of cases: the concepts were more general than we anticipated. For example, we realized that the mathematical notion of a topology maps incredibly well to the idea of physical distinguishability, but states of Hamiltonian system are not the only thing that are physically distinguishable.

Each new link often shrank the complexity of the arguments while expanding the scope of the work in a way that was not at all anticipated. The same link that allowed to justify Hamiltonian dynamics also gave us relativistic motion on the extended phase space. The same link that allowed to justify Lagrangian mechanics also constrained the motion to that of free particles under scalar/vector potential forces. We also realized that sometimes different mathematical concepts could be justified from the same physical starting point. That is: by putting physics first we could achieve a sort of unification of ideas that was not evident in the math alone.

@@

While trying to re-derive classical, quantum and (to some extent) relativistic mechanics may seem too broad a scope, it is necessary to show that the concepts we develop are truly general, that they can be applied to many different cases. If we could derive classical Hamiltonian mechanics on assumptions that are inapplicable to the quantum case, what good would it do? Quantum system are Hamiltonian systems, so clearly we wouldn't have understood what makes a system Hamiltonian. The ability to apply the same physical ideas far and wide is what gives us confidence that these are truly general concept that capture the physical aspects we are describing. The breath itself is what makes the work interesting and more than a mere curiosity.

@@

@@

The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology\cite{LeeTM}, measure theory (ref?), differential geometry (symplectic and Rimannian)\cite{Lee}, group theory and vector spaces\cite{Lang,Young}, statistics (ref?), information theory (ref?), Hamiltonian mechanics, Lagrangian mechanics\cite{classical_dynamics}, quantum mechanics (ref?), thermodynamics (ref?), special relativity (ref?) and so on. Using idea from all these areas is necessary, in fact unraveling their connections is part of the value of this work, but it does create a number of problems.

@@

This link between state definition and deterministic processes should not be too surprising as the state, in the context of control theory and dynamical systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.(ref?) As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can indeed be calculated, measured and prepared. We can also describe the evolution of each element of the distribution provided that: we have a way to isolate it and study it under deterministic and reversible motion (so that we can define microstates); the non-deterministic motion does not alter the system (the set of microstates is preserved by the evolution).

@@

However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy\cite{Shannon} is finite.} Note that different outcomes for the same process can overlap. For example, $4.12 \pm 0.05$ cm and $4.13 \pm 0.05$ cm are both legitimate possible outputs of the same measurement device. But since all states must be distinguishable, given two arbitrary states there must be a process precise enough to tell them apart. That is, the potential outcomes associated with the two states do not overlap. For example, $4.12 \pm 0.0005$ cm and $4.13 \pm 0.0005$ cm do not overlap anymore.

@@

Given that scientific reductionism (i.e. the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics,(ref?) we now explore how to characterize a system in terms of its components. That is, we want to study the relationship between the state space of a composite system and the state space of its parts. In general, this is quite a complicated thing to do, which requires intimate knowledge of the system at hand. So we simplify our problem and study a material made of infinitesimal homogeneous parts. What we'll find is that under the additional assumption that the material is infinitesimally reducible (i.e. its state is equivalent to the states of its infinitesimal parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.

@@

As we identified the space of Lebesgue integrable differentiable functions $C^1(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the space of distributions that are physically meaningful, we can better understand why other commonly used function spaces do not fit the bill. Some are not restrictive enough. The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{prop:continuity} are unphysical. The space of continuous functions $C(\mathcal{S})$, the space of continuous functions that vanish at infinity $C_0(\mathcal{S})$ and the space of differentiable functions $C^1(\mathcal{S})$ include functions whose integral is infinite, which would represent an infinite amount of material. Some definitions are too restrictive. Requiring compact support (i.e. the function is different from zero only on a finite region) would exclude distributions, such as Gaussians, that span over the whole range of states. Such distributions are physical: if we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range. Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible. The Whitney approximation theorem(ref?) makes this mathematically well defined.}

@@

But probably the easiest way to look at it is that initial and final sets of states have to possess the same count of states and possibilities as defined by the symplectic form $\omega$. Therefore the area within a degree of freedom, which represents the number of possibilities in said d.o.f., needs to be mapped to an equal area within the transported d.o.f. (i.e. the d.o.f. defined by the transported state variables). Independent d.o.f. must remain independent, and therefore transported orthogonal d.o.f. remain orthogonal. This means that the product of possibilities of independent d.o.f. is also conserved. These statements are the physical justification of Gromov's non-squeezing theorem \cite{Gromov,deGosson,Stewart} and Liouville's theorem. And they give intuitive insight on the geometry of Hamiltonian systems.

@@

The third case is when we perform a state variable transformation $\hat{q}=\hat{q}(q,t)$ that is time dependent (e.g. $\hat{q}=q+vt$). Our composite state distribution $\rho(q)$ becomes $\rho(\hat{q},t)$: it is no longer defined at equal time. This means that our measure $\mu$ needs to be modified to define integrals over the time variable.\footnote{Looking ahead, another way to frame the problem is that Galilean transformations do not preserve phase space volume. Lorentz transformations, on the other hand, do\cite{Treumann} and it is this requirements plus the kinematic assumption that will recover them later.}

@@

As we said before, the time variable is not simply another state variable: it doesn't identify extra configurations and it's the variable on which we have defined deterministic and reversible evolution. When changing variables, we still need to make sure that deterministic and reversible motion can be defined on the new time variable. Transformations like $\hat{t}=q, \; \hat{q}=t$ (exchanging time with another state variable) or $\hat{t}=\sqrt{t^2 + q^2}, \; \hat{q}=\tan^{-1}(t/q)$ (polar coordinates between time and state variable) clearly do not guarantee deterministic and reversible motion over the new time variable. To be meaningful, the change of time variable must preserve time ordering therefore the transformation has to be strictly monotonic between the two time variables. On the other hand, the change of state variables must preserve the ability to uniquely identify states at each instant in time. In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight.\cite{Gourgoulhon}}

@@

We recognize the equations as those for Hamiltonian mechanics on the extended phase space.\cite{Synge, Struckmeier}

@@

The link between kinematics, the study of motion through quantities like velocity and acceleration, and dynamics, the study of state evolution and its causes, is crucial in physics. In Newtonian physics, $\vec{F}=m\vec{a}$ provides such link. In Hamiltonian (and Lagrangian) mechanics, the link is implicitly given by the choice of the Hamiltonian (or Lagrangian).(ref?)

@@

This also allows us to understand conceptually why the speed of deterministic and reversible motion cannot be greater than $c$. Consider the trajectory of a point particle. Over an interval $dt$ it will travel $dx$ in space. Since the motion is deterministic and reversible, and therefore continuous, it will travel through all the states defined in the interval $dx$. It will also travel through every moment in time, which corresponds to $cdt$ possibilities. As it moves, time will always change but not necessarily space. The number of states traveled in space cannot exceed the number of states traveled in time. Therefore we have:
\begin{align*}
\frac{|n_x(dx)|}{|n_t(dt)|} &=  \frac{|dx|}{|c dt|} \leq 1 \\
\frac{|dx|}{|dt|} & \leq c \\
\end{align*}
Note, though, that this constraint applies only to deterministic and reversible motion. It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$. Non-local quantum effects, such as the ones featured in EPR\cite{EPR} experiments, are not examples of deterministic and reversible motion. Therefore they do not have anything to violate: the maximum speed does not apply to them. On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.

@@

Suppose we have our classical electron described by a finite amount of classical material distributed over a single d.o.f. $\int \rho_\mathcal{c} dq \wedge dp = 1$ where $\rho_\mathcal{c}$ is the normalized distribution. In this context, we will call \emph{fragment} an infinitesimal part of our classical electron. The Shannon entropy\cite{Shannon} associated with the distribution is finite as well. $I(\rho_\mathcal{c}) = - \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$. Intuitively, this value can be thought of as the amount of information required to identify a fragment within the distribution.\footnote{To be precise, Shannon's differential entropy is actually the number of bits required to identify an element of the given distribution relative to one of a uniform distribution of unitary range on all axes.}

@@

	This notion may seem in contradiction to the current practice. For example, the S-matrix, one of the main devices for calculating scattering amplitudes and cross sections, is derived by assuming unitary evolution.(ref?) Doesn't it imply we are assuming that deterministic and reversible evolution is happening at every instant?

@@

@@

\bibitem{Shannon} Shannon, C. E., ``A mathematical theory of communication'', The Bell System Technical Journal, Vol. 27, pp. 379--423, 623--656, (1948).
\bibitem{Jaynes} Jaynes, E. T., ``Information theory and statistical mechanics'', Statistical Physics 3, pp. 181--218, (1963).
\bibitem{classical_dynamics} J. V. Jos\'{e}, E. J. Saletan, ``Classical Dynamics'', Cambridge University Press, (1998).
\bibitem{Gromov} Gromov, M. L., ``Pseudo holomorphic curves in symplectic manifolds'', Inventiones Mathematicae 82, pp. 307--347, (1985).
\bibitem{deGosson} de Gosson, M. A., ``The symplectic camel and the uncertainty principle: the tip of an iceberg?'', Foundations of Physics 39, pp. 194--214, (2009).
\bibitem{Stewart} Stewart, I., ``The symplectic camel'', Nature 329, pp. 17--18, (1987).
\bibitem{Lanczos} Lanczos, C., ``The variational principles of mechanics'', University of Toronto Press, (1949).
\bibitem{Synge} Synge, J. L., ``Classical dynamics'', Encyclopedia of Physics Vol 3/1, Springer (1960).
\bibitem{Struckmeier} Struckmeier, J., ``Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems'', J. Phys. A: Math. Gen. 38, pp. 1257--1278, (2005).
\bibitem{Lee} Lee, J. M.: Introduction to Smooth Manifolds. Springer (2003)
\bibitem{LeeTM} Lee, J. M.: Introduction to Topological Manifolds. Springer (2000)
\bibitem{Lang} Lang, S. L.: Algebra. Springer (2002)
\bibitem{Young} Young, N.: An Introduction to Hilbert Space. Cambridge Univ. Press, Cambridge, UK (1988)

\bibitem{Barron} Barron, R. F.: 8.2 Lumped-parameter analysis. In: Barron, R. F.: Industrial Noise Control and Acoustics. CRC Press (2002)
%online at http://ssu.ac.ir/cms/fileadmin/user_upload/Daneshkadaha/dbehdasht/behdasht_herfei/e_book/_Randall_F._Barons_Industrial_Noise_Control_and_.org__1_.pdf

\bibitem{Roland}  
 Thomas, R. E., Rosa, A. J.: The analysis and design of linear circuits. Wiley (2001)

\bibitem{Borutzky} Borutzky, W.: Bond graph methodology: development and analysis of multidisciplinary dynamic system models.
Springer (2010)

\bibitem{Freiding} Freidin, R., Lasnik, H.: Syntax critical concepts in linguistics. Routledge (2006)

\bibitem{Friedman} Friedman, D. P., Wand, M., Haynes, C.T.: Essentials of Programming Languages. MIT Press (2001)

\bibitem{Treumann} Treumann, R. A., Nakamura, R., Baumjohann,  W.: Relativistic transformation of phase-space distributions. Annales Geophysicae (2011)

\bibitem{Gourgoulhon} Gourgoulhon, E.: 3+1 Formalism and Bases of Numerical Relativity. arXiv. http://arxiv.org/abs/gr-qc/0703035 (2007)

\bibitem{EPR} Einstein, A., Podolsky, B.,  Rosen, N.: Can quantum-mechanical description of physical reality be considered complete? Phys. Rev. 47 777 (1935)
